---
title: "Hypothesis Testing, Part 1"
author: "Nicholas A. Del Grosso"
date: "July 23, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
x <- seq(-6, 12, .01)
y1 <- dnorm(x, mean=-1, sd=1,)
y2 <- dnorm(x, mean=4, sd=2)
xlim <- c(-4, 10)
ylim <- c(0, .5)
plot(x, y1, 'l', ylab='', col=rgb(1,0,0,1), main="Two Normal Population Distributions, with Sample Data (N=100)", xlim=xlim, ylim=ylim)
lines(x, y2, col=rgb(0,0,1,1))
par(new=TRUE)
hist(rnorm(100, mean=-1, sd=1), 12, xlim = xlim, col=rgb(1,0,0,0.2), freq=F, ylim=ylim, main='', xlab='')
par(new=TRUE)
hist(rnorm(100, mean=4, sd=2), 12, xlim = xlim, col=rgb(0,0,1,0.2), freq=F, ylim=ylim, main='', xlab='')
```

## Querying and Generating Probability Densities

R's built-in stats package comes with lots of functions for working with common probability distributions.  The functions we will review here allow you to:

  - Get the distribution's **(d)ensity** height, given x values: *dnorm()*, *dpois()*, *dunif()*
  - Get **(r)andom** amples from the distribution: *rnorm()*, *rpois()*, *runif()*
  - Get the distributions's quantiles: 
    - The **(q)uantile** of the distribution, given a percentile: *qnorm()*, *qpois()*, *qunif()*
    - The **(p)ercentile** of the distribution, given a quantile: *pnorm()*, *ppois()*, *punif()*
    
Examples of usage:

```{r}
x <- seq(-3, 3, .1)
y <- dnorm(x, mean=0, sd=1)
plot(x, y)

data <- rnorm(100, mean=0, sd=1)
hist(data)

qnorm(p=.05, mean=0, sd=1)
pnorm(q=-1.645, mean=0, sd=1)

```


## T-Tests

The **t.test()** function does t-tests.  Setting different options in the function arguments lets you use it as one-sided or two-sided test, a paired-sample or independent-sample test, assume equal variance, etc.  It returns a table of the results:

```{r}
x <- rnorm(100, mean=-0.5, sd=1)
y <- rnorm(100, mean=0.5, sd=1)
t.test(x, y, alternative = "two.sided")
```

Nonparametric alternatives to the t-test are also available:
  - Kruskall-Wallis Test: *kruskall.test()*
  - Mann-Whitney U Test: *wilcox.test()*
  - Wilcoxon Rank Sum Test: *wilcox.test(paired=TRUE)*

  
If you save the result of the test as a variable, you can also access the details of the table and index the **names** of the result to access the data:

```{r}
results = t.test(x, y, alternative = "two.sided")
names(results)

results['p.value']  # Can also be written as: results$p.value
```

## Correlations

Testing for correlations between paired samples can be done with the *cor.test()* function.  By default it does a Pearson correlation, but can also do a Spearman or Kendall's Tau test by changing the "method" argument



## Exercises

Exercises are available at the hypothesis_testing_exercises.R script.



